On Devon Allen's Disqualification at the 2022 Track and Field World Championship

### 2023-04-25 Manuscript/manuscript.tex
Started to make changes

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{red}{EDS: (#1)}}

@@

%1)
Devon Allen, \eds{a student (or an alumnus)?} of the University of Oregon in 
Eugene, Oregon, was
expected to be a contender for the 2022 World Track and Field Championships 110 
meter hurdle.  After running the event in 12.84 seconds in early June 2022 
(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
in third in the event at the United States Track and Field 
Championships also in Eugene at the end of June, Allen was poised to 
compete in the 2022 World Track and Field Championships again in Eugene in August.
%After coming in third in the 110 meter hurdles at the United States Track and 
%Field Championships also in Eugene at the end of June, Allen was poised to 
%compete in the 2022 World Track and Field Championships also 
%located at Eugene in August.
%two months later. 
At the World Championships, Allen made it through the 
heats and semifinals to reach the finals while posting reaction times of 0.123
in the heats \eds{if more than one heat, include all of them or the average;
if only one heat, change to singular form} and 0.101 in the semifinal.  
%Allen, who attended the University of Oregon also in Eugene, 
%(the location of US World Championship 
%\eds{US World? remove World?} and the World Championship) 
%was
%considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
%in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
%Thus 
His disqualification by reacting in 0.099 seconds instead of 0.1 was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that showed the disqualification; it can be found here:
\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
when they found out that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete and been 
spared waiting another two years until the next World Track and Field Championship.
An investigation into reaction times is also important because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times; which is the goal of
every sprinter who steps onto the track.
\eds{final two sentence seems out of place here; consider moving elsewhere?}

@@

%4)
\jy{language needs to be formalized and condensed.}
The objective to this paper is to analyze the reaction times at the 2022
World Track and Field Championship to check for any abnormalities, and we approach
this two different ways.  Our first method is to try to model the data, which is
the reaction times of every non-disqualification since the 1999 World Championships. 
The model we use is a gamma linear mixed-effects model with a random effect
accounting for the venue (or year) effect. 
The second method is to compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship. 
%\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
%think it is critical for readers to know all the details here.}
%First we can compare the reaction times of athletes who
%competed at the World Track and Field Championship in both 2019 and 2022, and 
%second we can compare how athletes performed at the 2022 championship relative 
%to their national/qualifiers meet. In order to reduce any chances
%of athletes improving reaction time, we can also look at data from qualifying meets.
%Almost every athlete had to qualify by proving they are elite at the national level, 
%usually by coming in at least fourth. Thus, data was compiled on athletes who competed
%at the 2022 World Track and Field Championships on what these athletes also ran 
%at their country's national championship. If there was any argument that athletes
%have improved their reaction time from 2019 to 2022, it should now be considered
%nearly null, as the national competitions took place between June and July 2022 
%and the World Championships were in August. The equipment used to measure reaction
%time cannot be guaranteed to be the same from national to international
%competition, but the time span between meets has been reduced considerably.

>>>>

\newcommand{\jy}[1]{\textcolor{blue}{JY: #1}}
\newcommand{\eds}[1]{\textcolor{red}{EDS: (#1)}}
\newcommand{\of}[1]{\textcolor{violet}{OF: #1}}

@@

%1)
Devon Allen, an alumnus of the University of Oregon in 
Eugene, Oregon, was
expected to be a contender for the 2022 World Track and Field Championships 110 
meter hurdle.  After running the event in 12.84 seconds in early June 2022 
(only 0.04 seconds away from the world record \citep{wa2022preview}) and coming 
in third in the event at the United States Track and Field 
Championships also in Eugene at the end of June, Allen was poised to 
compete in the 2022 World Track and Field Championships again in Eugene in August.
%After coming in third in the 110 meter hurdles at the United States Track and 
%Field Championships also in Eugene at the end of June, Allen was poised to 
%compete in the 2022 World Track and Field Championships also 
%located at Eugene in August.
%two months later. 
At the World Championships, Allen made it through the 
heats and semifinals to reach the finals while posting reaction times of 0.123
in the heats \eds{if more than one heat, include all of them or the average;
if only one heat, change to singular form}
\of{There are roughly 5 heats, but each athlete only competes in one heat} 
and 0.101 in the semifinal.  
%Allen, who attended the University of Oregon also in Eugene, 
%(the location of US World Championship 
%\eds{US World? remove World?} and the World Championship) 
%was
%considered to be a favorite for the 110 meter hurdle after running 12.84 seconds
%in early June; only 0.04 seconds away from the world record \citep{wa2022preview}.
%Thus 
His disqualification by reacting in 0.099 seconds instead of 0.1 was met
with shock from both Allen and fans alike.  NBC, who was broadcasting the
championships in the United States, later uploaded a video to their YouTube
channel that showed the disqualification; it can be found here:
\url{https://www.youtube.com/watch?v=D6NXTMo-1yM}. The crowd erupted with jeers 
when they found out that Allen was disqualified, angry at the result.  Had Allen 
reacted just 0.001 seconds slower he would have been allowed to compete and been 
spared waiting another two years until the next World Track and Field Championship.

\eds{final two sentence seems out of place here; consider moving elsewhere?}
\of{Moved to the end of paragraph 4 of introduction}

@@

%4)
\jy{language needs to be formalized and condensed.}
The objective to this paper is to analyze the reaction times at the 2022
World Track and Field Championship to check for any abnormalities, and we approach
this two different ways.  First we try to model the data, which is
the reaction times of every non-disqualification since the 1999 World Championships. 
The model we use is a gamma linear mixed-effects model with a random effect
accounting for the venue (or year) effect. 
The second method is to compare reaction times for athletes who have competed
at additional competitions to the 2022 World Track and Field Championship. 
Our investigation into reaction times is important because reaction times have
been shown to affect the overall time of the race \citep{delalija2008reaction}.
Thus, faster reaction times lead to faster overall times; which is the goal of
every sprinter who steps onto the track.
%\eds{I recommend moving the rest of this paragraph to Section 2.2 as I don't
%think it is critical for readers to know all the details here.}
%First we can compare the reaction times of athletes who
%competed at the World Track and Field Championship in both 2019 and 2022, and 
%second we can compare how athletes performed at the 2022 championship relative 
%to their national/qualifiers meet. In order to reduce any chances
%of athletes improving reaction time, we can also look at data from qualifying meets.
%Almost every athlete had to qualify by proving they are elite at the national level, 
%usually by coming in at least fourth. Thus, data was compiled on athletes who competed
%at the 2022 World Track and Field Championships on what these athletes also ran 
%at their country's national championship. If there was any argument that athletes
%have improved their reaction time from 2019 to 2022, it should now be considered
%nearly null, as the national competitions took place between June and July 2022 
%and the World Championships were in August. The equipment used to measure reaction
%time cannot be guaranteed to be the same from national to international
%competition, but the time span between meets has been reduced considerably.
